Anony\emph{mice}d shareable data: Using \emph{mice} to create and analyze multiply imputed synthetic data sets

### 2021-10-26 manuscript/Manuscript.tex
Start revising paper.

@@

Open science, including open data, has been marked as the future of
science \citep{gewin_data_2016}, and the advantages of publicly
available research data are numerous
\citep{molloy_open_2011, walport_brest_sharing_2011}. Collecting
research data requires an enormous investment both in terms of time and
monetary resources. Openly accessible research data bears the potential
of increasing the scientific returns for the same data collection
effort. Additionally, the fact that public funds are used for data
collection results in increasing demand for access to the collected
data. Nevertheless, the possibilities to distribute research data
directly are often very limited due to restrictions on data privacy and
data confidentiality. Although these regulations are much needed,
privacy constraints are also ranked among the toughest challenges to
overcome in the advancement of modern day social science research
\citep{lazer_life_2009}.

@@

Conceptually, the synthetic data framework is based upon the building
blocks of multiple imputation of missing data, as proposed by
\citet{rubin_multiple_1987}. Instead of replacing just the missing
values with multiple draws from the posterior predictive distribution,
one could easily \emph{overimpute} any observed sensitive values.
Similarly to multiple imputation of missing data, the multiple synthetic
data sets allow for correct statistical inferences, despite the fact
that the analyses do not use the ``true'' value. The analyses over
multiple synthetic data sets should be pooled into a single inference,
so that the researcher can draw valid conclusions from the pooled
results. To that respect, the variance should reflect the added
variability that is induced by the imputation procedure.

@@

To mitigate unnecessary challenges related to creating synthetic data
sets on behalf of the researcher, software aimed at multiple imputation
of missing data can be employed. Especially if researchers acquired
familiarity with this software during earlier projects, or used it
earlier during the research process, the additional burden of creating
synthetic data sets is relatively small. The R-package \texttt{mice}
\citep{mice} implements multiple imputation of missing data in a
straightforward and user-friendly manner. However, the functionality of
\texttt{mice} is not restricted to the imputation of missing data, but
allows to impute any value in the data: even observed values.
Consequently, \texttt{mice} can be utilized for the creation of multiply
imputed synthetic data sets.

@@

In \texttt{mice}, the generation of multiply imputed data sets to solve
for unobserved values is straightforward. The following pseudocode
details the multiple imputation of the \texttt{mice::boys} data set
\citep{fredriks_boys_2000} into the object \texttt{imp} with
\texttt{m\ =\ 10} imputated sets and \texttt{maxit\ =\ 7} iterations for
the algorithm to converge, using the default imputations methods for
each column data class.

@@

It is straightforward to extend the imputation approach to generate
synthetic values. Rather than imputing missing data, the observed values
are then replaced by synthetic draws from the posterior predictive
distribution. For simplicity, assume that the data is completely
observed (i.e., \(\textbf{Y} = \textbf{Y}_{obs}\)). Following the
notation of \citet{reiter_raghunathan_multiple_2007}, let for \(n\)
units denote \(Z_i = 1\) if any of the values of unit
\(i = 1, 2, \dots, n\) are to be replaced by imputations, and
\(Z_i = 0\) otherwise, with \(\textbf{Z} = (Z_1, Z_2, \dots, Z_n)\).
Accordingly, the data consists of values that are to be replaced and
values that are to be kept (i.e.,
\(\textbf{Y} = (\textbf{Y}_{rep}, \textbf{Y}_{nrep})\). Now, instead of
imputing \(\textbf{Y}_{mis}\) with draws from the posterior predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) as in the
missing data case, \(\textbf{Y}_{rep}\) is imputed from the posterior
distribution of
\(P(Y^{(l)}_{j, rep} | \textbf{Y}^{(l)}_{-j}, \textbf{Z}, \theta)\),
where \(l\) is an indicator for the synthetic data set
(\(l = 1, 2, \dots, m\)). Note that synthetic values that are imputed at
an earlier step can be used for imputing variable \(j\). This process
results in the synthetic data
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\).

For example, overimputing synthetic values for both the observed and
missing cells in the \texttt{mice::boys} data set into the object
\texttt{syn}, given the same imputation parameters as before, can be
realized by the following code execution.

@@

where the argument \texttt{where} requires a matrix of the same
dimensions as the data, (i.e., a \(n \times k\) matrix) containing
logicals \(z_{ij}\) that indicate which cells are selected to have their
values replaced by draws from the posterior predictive distribution.
This approach allows to overimpute a subset of the observed data, or as
in the above example, the observed data as a whole, resulting in a data
set that partially or completely consists of synthetic data values.

@@

Also similarly to the missing data case, variances, and subsequently
confidence intervals, should incorporate the increase in variance that
is due to imputation
\citep{reiter_partially_inference_2003, drechsler_synthetic_2011}. Yet,
the increase in variance due to imputation differs according to whether
missing values are imputed or observed data is overimputed with missing
values. Whereas the variance estimate after imputation of missing data
needs to account for the fact that a certain amount of information in
the data is missing, variance estimation from synthetic data does not
suffer from this issue. The adjusted variance estimate that follows from
using multiple synthetic data sets only suffers from the fact that a
finite number of \(m\) synthetic data sets are used to resemble the
observed data. Hence, the according variance estimate for synthetic data
as developed by \citet{reiter_partially_inference_2003} yields \[
T = \bar{u}_m + \frac{b_m}{m},
\]

@@

Additionally, if not all variables in the data are synthesized, but the
variables that are synthesized can be linked to open access data, a
disclosure risk may remain \citep{drechsler2011empirical}. If the open
access data allows for identification, and the corresponding
observations in the synthetic data can be identified, the variables that
are not synthesized may provide information that should not have been
disseminated. The associated problems generally decrease when the
sensitive variables are synthesized as well. Still, it is important to
remain critical regarding the extent to which the synthetic data might
be disclosive. The practical development of easy to use software to
identify which observations are at risk of disclosure is an area of
future work that can improve these issues. Subsequently, the
implementation of ways to overcome such problems, for example by record
swapping as suggested by \citet{jiang2021balancing}, once detected is
welcomed.

That said, while the fields of data disclosure control and data
synthetisation originated decades ago, it is more relevant today than
ever. In our study we outline a standard research workflow where the
goal is to synthesize a complete data set. To simplify the simulation we
adopted a bootstrapping scheme to induce sampling variance in order to
conform to the combination procedure defined by
\citet{reiter_partially_inference_2003}. Ideally, one would like to omit
the bootstrap from the synthetization to adopt a scheme where the
sampled data itself serves as a reference: much like the procedure
outlined in \citet{vink2014pooling} for incomplete data simulation. The
corresponding pooling rules have not been derived yet and the
incorporation in data analysis workflows would require proper attention
from developers alike.

It is important to note that in our simulations we used a single
iteration. A single iteration is sufficient only when the true data is
completely observed, or when the missingness pattern is monotone
\citep{drechsler_synthetic_2011}. If both observed and unobserved values
are to be synthesized, then more iterations and a careful investigation
into the convergence of the algorithm are in order. Synthetic data
generation with \texttt{mice} is therein no different than multiple
imputation with \texttt{mice}.

There are other developments that can generate synthetic data sets. For
example, the \texttt{synthpop} \citep{synthpop} package in \texttt{R}
would yield valid synthetic data sets, but requires the true data to be
completely observed. To avoid this, one could adopt a two-step approach
wherein the incomplete values are multiply imputed before
synthetization. Given \(m\) multiple imputations and \(r\)
synthetizations, at least \(m \times r\) synthetic data sets are then in
order. However, this approach requires that researchers first impute the
missingness using a different package, and then use \texttt{synthpop} to
create synthetic data sets, while \texttt{mice} allows to do both.
Additionally, the flexibility with \texttt{mice} is that both unobserved
and observed data values could be synthesized at once, without the need
for a two-step approach. Then, using \texttt{mice}, \(m\) synthetic sets
are sufficient. As of today, no pooling rules for one-step imputation of
missingness and synthetisation have been developed, but the derivation
of those would further reduce the burden of creating synthetic data
sets.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{We are grateful to Mirthe Hendriks and Stijn van den
Broek for replicating the simulation study on an independent data set.}

>>>>

\hypertarget{add-reviewer-name}{%
\section{ADD REVIEWER NAME}\label{add-reviewer-name}}

@@

Open science, including open data, has been marked as the future of
science \citep{gewin_data_2016}, and the advantages of publicly
available research data are numerous
\citep{molloy_open_2011, walport_brest_sharing_2011}. Collecting
research data requires an enormous investment both in terms of time and
monetary resources. Disseminating the research data bears the potential
of increasing the scientific returns for the same data collection effort
tremendously. Additionally, the fact that public funds are used for data
collection results in increasing demand for access to the collected
data. Nevertheless, the possibilities to distribute research data
directly are often very limited due to restrictions on data privacy and
data confidentiality. Although these regulations are much needed,
privacy constraints are also ranked among the toughest challenges to
overcome in the advancement of modern day social science research
\citep{lazer_life_2009}.

@@

Conceptually, the synthetic data framework is based upon the building
blocks of multiple imputation of missing data, as proposed by
\citet{rubin_multiple_1987}. Instead of replacing just the missing
values with multiple draws from the posterior predictive distribution,
one could easily replace any observed sensitive values. Similarly to
multiple imputation of missing data, the multiple synthetic data sets
allow for correct statistical inferences, despite the fact that the
analyses do not use the ``true'' value. The analyses over multiple
synthetic data sets should be pooled into a single inference, so that
the researcher can draw valid conclusions from the pooled results. To
that respect, the variance should reflect the added variability that is
induced by the imputation procedure.

@@

To mitigate unnecessary challenges related to creating synthetic data
sets on behalf of the researcher, software aimed at multiple imputation
of missing data can be employed. Especially if researchers acquired
familiarity with this software during earlier projects, or used it
earlier during the research process, the additional burden of creating
synthetic data sets is relatively small. The R-package \texttt{mice}
\citep{mice} implements multiple imputation of missing data in a
straightforward and user-friendly manner. However, the functionality of
\texttt{mice} is not restricted to the imputation of missing data, but
allows to impute any value in the data: even observed values.
Consequently, \texttt{mice} can be used for the creation of multiply
imputed synthetic data sets.

@@

In \texttt{mice}, the generation of multiply imputed data sets to solve
for unobserved values is straightforward. The following pseudocode
details the multiple imputation of the \texttt{mice::boys} data set
\citep{fredriks_boys_2000} into the object \texttt{imp} with
\texttt{m\ =\ 10} imputated sets and \texttt{maxit\ =\ 7} iterations for
the algorithm to converge, using the default imputations methods given
each column's data class.

@@

It is straightforward to extend the imputation approach to generate
synthetic values. Rather than imputing missing data, the observed values
are then replaced by synthetic draws from the posterior predictive
distribution. For simplicity, assume that the data is completely
observed (i.e., \(\textbf{Y} = \textbf{Y}_{obs}\)). Following the
notation of \citet{reiter_raghunathan_multiple_2007}, let, given \(n\)
observations, \(Z_i = 1\) if any of the values of unit
\(i = 1, 2, \dots, n\), are to be replaced by imputations, and
\(Z_i = 0\) otherwise, with \(\textbf{Z} = (Z_1, Z_2, \dots, Z_n)\).
Accordingly, the data consists of values that are to be replaced and
values that are to be kept (i.e.,
\(\textbf{Y} = (\textbf{Y}_{rep}, \textbf{Y}_{nrep})\). Now, instead of
imputing \(\textbf{Y}_{mis}\) with draws from the posterior predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) as in the
missing data case, \(\textbf{Y}_{rep}\) is imputed from the posterior
distribution of
\(P(Y^{(l)}_{j, rep} | \textbf{Y}^{(l)}_{-j}, \textbf{Z}, \theta)\),
where \(l\) is an indicator for the synthetic data set
(\(l = 1, 2, \dots, m\)). The synthetic values that are imputed at an
earlier step can be used for imputing variable \(j\). This process
results in the synthetic data
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\).
Note that, given a completely observed data set, iterating over the
posterior predictive distribution is not required.

If the goal is to create synthetic versions of a data set that suffers
from missingness, the imputation procedure becomes more complex. A
two-step imputation procedure has been proposed by
\citet{reiter2004simultaneous}. In the first step, the missing data is
imputed according to the principles of multiple imputation for missing
data, while in a second step, the synthetic data is imputed conform the
corresponding procedure. Intuitively, these steps could be combined
using \texttt{mice}, conditioning the imputations on both the
missingness and the values that are to be synthesized. In practice,
replacing observed and missing cells in the \texttt{mice::boys} data set
by synthetic values into the object \texttt{syn}, given the same
imputation parameters as before, can be realized by the following code
execution.

@@

Here, the argument \texttt{where} requires a matrix of the same
dimensions as the data, (i.e., a \(n \times k\) matrix) containing
logicals \(z_{ij}\) that indicate which cells are selected to have their
values replaced by draws from the posterior predictive distribution.
This approach allows to replace a subset of the observed data, or as in
the aforementioned example, the observed data as a whole, resulting in a
data set that partially or completely consists of synthetic data values.
However, because the validity of this approach when the data suffers
from missingness has not been investigated yet, we restrict ourselves in
the following sections to a completely observed data set.However,
because the validity of this approach has not been investigated yet, we
restrict ourselves in the followi

@@

Also similarly to the missing data case, variances, and subsequently
confidence intervals, should incorporate the increase in variance that
is due to imputation
\citep{reiter_partially_inference_2003, drechsler_synthetic_2011}. Yet,
the increase in variance due to imputation differs according to whether
missing values are imputed or observed data is replaced by synthetic
values. Whereas the variance estimate after imputation of missing data
needs to account for the fact that a certain amount of information in
the data is missing, variance estimation from synthetic data does not
suffer from this issue. The adjusted variance estimate that follows from
using multiple synthetic data sets only suffers from the fact that a
finite number of \(m\) synthetic data sets are used to resemble the
observed data. Hence, the according variance estimate for synthetic data
as developed by \citet{reiter_partially_inference_2003} yields \[
T = \bar{u}_m + \frac{b_m}{m},
\]

@@

Additionally, if not all variables in the data are synthesized, but the
variables that are synthesized can be linked to open access data, a
disclosure risk may remain \citep{drechsler2011empirical}. If the open
access data allows for identification, and the corresponding
observations in the synthetic data can be identified, the variables that
are not synthesized may provide information that should not have been
disseminated. The associated problems generally decrease when the
sensitive variables are synthesized as well. Still, it is important to
remain critical regarding the extent to which the synthetic data might
be disclosive. The practical development of easy to use software to
identify which observations are at risk of disclosure is an area of
future work that can improve these issues. Subsequently, the
implementation of ways to overcome such problems once detected, for
example by record swapping as suggested by \citet{jiang2021balancing},
is welcomed.

That said, while the fields of data disclosure control and data
synthetisation originated decades ago, it is more relevant today than
ever. In our study we outline a standard research workflow where the
goal is to synthesize a complete data set. To simplify the simulation we
adopted a bootstrapping scheme to induce sampling variance in order to
conform to the combination procedure defined by
\citet{reiter_partially_inference_2003}. Ideally, one would like to omit
the bootstrap from the synthetisation to adopt a scheme where the
sampled data itself serves as a reference: much like the procedure
outlined in \citet{vink2014pooling} for incomplete data simulation. The
corresponding pooling rules have not been derived yet and the
incorporation in data analysis workflows would require proper attention
from developers alike.

There exist other software that is capable of creating synthetic data.
For example, the \texttt{synthpop} \citep{synthpop} package in
\texttt{R} pioneered in the field of synthetic data generation in
\texttt{R}, and contains many functionalities to create and evaluate
synthetic data sets. However, although \texttt{synthpop} is capable to
synthesize regardless of missing data, it cannot solve eventual
missingness problems. If the data for synthetisation contains missing
values, the missingness will be considered as a property of the data,
and the corresponding synthetic data sets will contain missing values as
well. If the goal is to impute and synthesize a data set, a two-step
approach must be used, in which the missing values are imputed using a
different package, like \texttt{mice}, after which \texttt{synthpop} can
be used to synthesize the imputed data. Then, given \(m\) multiple
imputations and \(r\) synthetisations, at least \(m \times r\) synthetic
data sets are in order, after which the data can be analysed and the
analyses can be pooled using rules developed by
\citet{reiter2004simultaneous}. However, such an approach would be
computationally inefficient and practically cumbersome, because the
architectures of both packages differ.

The appealing nature of \texttt{mice} therefore lays in its ability to
solve both the missingness problem and create synthetic versions of the
data. Eventually, the flexibility with \texttt{mice} is that both
unobserved and observed data values could be synthesized at once,
without the need for a two-step approach. Using \texttt{mice}, \(m\)
synthetic sets would sufficient. However, as of today, no pooling rules
for one-step imputation of missingness and synthetisation have been
developed. The derivation of those would yield a massive theoretical
improve to further reduce the burden of creating synthetic data sets.
More practically, the \texttt{synthpop} package contains functionality
that would be highly beneficial to users in \texttt{mice}, while
\texttt{mice} is likely to contain functionality that would be
beneficial to \texttt{synthpop}. Although efforts have been made to
bridge gaps between the packages \citep[e.g.,][]{miceadds_3.11-6},
further improvements could be made in that respect. Specifically,
\texttt{synthpop} contains a wide variety of functions that can be used
to assess the quality of synthetic data {[}e.g., see ADD SNOKE REFERENCE
+ ARXIV{]}, while \texttt{mice} contains methods to impute multilevel
data. Being able to use such functionality interchangeably would greatly
benefit the public.

It is important to note that in our simulations we used a single
iteration. A single iteration is sufficient only when the true data is
completely observed, or when the missingness pattern is monotone
\citep{drechsler_synthetic_2011}. If both observed and unobserved values
are to be synthesized, then more iterations and a careful investigation
into the convergence of the algorithm are in order. Synthetic data
generation with \texttt{mice} is therein no different than multiple
imputation.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{We are grateful to Mirthe Hendriks and Stijn van den
Broek for replicating the simulation study on an independent data set.
Additionally, we thank Hanne Oberman, {[}{[}NOT ANONYMOUS REVIEWER{]}{]}
and an anonymous reviewer for their stimulating comments.}
